---
title: "Coursera Practical Machine Learning"
author: "Karthik"
date: "10/21/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.  

The goal is to predict Human Acticity Recognition for weight lifting excercise based on the data collected sensors placed on arm, belt, forearm and dumbbell.
The excerise performed was Barbel lifts in 5 different ways
Class A - correct way of preforming barbell lifts
Class B - Throwing the elbows to the front
Class C - Lifting the dumbbell only halfway 
Class D - Lowering the dumbbell only halfway
Class E - Throwing the hips to the front

Thanks to the autors (Ugulino,W.; Cardador,D.; Vega,K.; Velloso,E.; Milidiu, R.; Fuks, H.) of the publication [WearableComputing: Accelerometers' Data Classification of Body Posturesand Movements](http://groupware.les.inf.puc-rio.br/work.jsf?p1=10335) for generously allowing the data collected to be used for this project 

# Environment Setup
```{r EnvironmentSetup, message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```
# Data Retrieval & Processing
```{r Data}
trainRawData <- read.csv("pml-training.csv")
testRawData <- read.csv("pml-testing.csv")

#The training data set contains 19622 observations and 160 variables, and the testing data set contains 20 observations and 160 variables.The "classe" variable in the training set is the outcome to predict.
dim(trainRawData)
dim(testRawData)

sum(complete.cases(trainRawData))

#Data cleaning - remove Columns with NA and remove unwanted timestamp information 
trainRawData <- trainRawData[, colSums(is.na(trainRawData)) == 0] 
testRawData <- testRawData[, colSums(is.na(testRawData)) == 0] 

trainRawData$classe = factor(trainRawData$classe)
classe <- trainRawData$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRawData))
trainRawData <- trainRawData[, !trainRemove]
trainCleaned <- trainRawData[, sapply(trainRawData, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRawData))
testRawData <- testRawData[, !testRemove]
testCleaned <- testRawData[, sapply(testRawData, is.numeric)]
# set seed producing repetative output
set.seed(1123)
#Split the training data set into 70% training data set and 30% for cross validation</p>
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

# Model
Random Forest Algorithm is used as predictive model for activity recognition because it can handle large number of inputs when interaction between variables is unknown, its accuracy and it automatically picks important variables
5 fold cross validation is used in the algorithm

# Method: Random Forest
```{r RandomForest}
control <- trainControl(method="cv", 5)
model <- train(classe ~ ., data=trainData, method="rf", trControl=control, ntree=250)
model
predictTest <- predict(model, testData)
confusionMatrix(testData$classe, predictTest)
accuracy <- postResample(predictTest, testData$classe)
accuracy
sampleError <- 1 - as.numeric(confusionMatrix(testData$classe, predictTest)$overall[1])
sampleError
```
# The model had an accuracy of 99.05% and out of sample error is 0.95%



# Prediction
Predict Excercise activity using the Model

```{r Prediction}
result <- predict(model, testCleaned[, -length(names(testCleaned))])
result

```
